---
title: "Lastfm_Clustering"
author: "Pranay Singla"
date: "1/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The aim of this code file is to cluster users from lastfm dataset into groups based on listening preferences. This analysis can be used to identify users with similar listerning preferences as well as their artist/genre preferences.

```{r}
# reading data
lastfm = read.csv("lastfm1.csv")
```

Most popular artist

```{r}
which.max(colSums(lastfm[,-1]))
```

Artist most correlated with Beatles

```{r}
beatles_index = grep('the.beatles', colnames(lastfm))
temp.cors = cor(lastfm$the.beatles, lastfm[,-c(1,beatles_index)])

# See which artist this is:
names(lastfm[,-c(1,465)])[which.max(temp.cors)]

```

Running hierarchical clustering to group users

```{r}
dist.mat = dist(lastfm[,-1], method = "euclidean")
lastfm.clust = hclust(dist.mat, method = "ward.D")
```

Plotting the dendogram

```{r}
# Dendrogram
plot(lastfm.clust)
```

Creating 5 clusters based on dendogram and listing number of users in each cluster

```{r}
# Create the clusters:
clust.groups = cutree(lastfm.clust, k = 5)
table(clust.groups)

```

Number of users in each cluster:  
1 -> 3798  
2 -> 3262  
3 -> 6037  
4 -> 1973  
5 -> 1089  


  
Finding top 6 artists in different clusters
```{r}

# Display top 6 artists in cluster 3:
print('Top 6 artists in Cluster 3')
tail(sort( colMeans(lastfm[clust.groups == 3,-1]), decreasing = F ), 6)

```

Users in cluster 3 prefer to listen to classic rock artists

```{r}
# Display top 6 artists in cluster 4:
print('Top 6 artists in Cluster 4')
tail(sort( colMeans(lastfm[clust.groups == 4,-1]), decreasing = F ), 6)
```

Users in cluster 4 prefer to listen to pop artists

```{r}
# Display top 6 artists in cluster 5:
print('Top 6 artists in Cluster 5')
tail(sort( colMeans(lastfm[clust.groups == 5,-1]), decreasing = F ), 6)
```

Users in cluster 5 prefer to listen to rap artists